---
title: "Data Simulation for Multivariate Linear Models"
author: "Erick A. Chacon-Montalvan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Simulation for Linear Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
knitr::opts_chunk$set(comment = "#>")
options(width = 100)
```

In this vignette, we show how to use the three main functions of the package
`datasim` which are `sim_data`, `model_frame` and `model_response`. This is an
introductory tutorial and only the simulation of linear Gaussian models are
presented.

#### Required packages

```{r}
library(datasim)
```

## Linear Model

First, we need to define a list of formulas specifying the type of effect that are
included in the linear predictor of each parameter. For example, this list can be
defined as follows.

```{r}
f <- list(
  mean ~ mfe(x, beta = c(-10, 0, 10)),
  sd ~ mfa(ones, beta = matrix(c(0.1, 0.5, 3), nrow = 1))
  )
```

In this formula, it can be seen that an intercept, a linear effect on `x1` and a
factor effect on `sex` are being included on the `mean` parameter, while the
standard deviation `sd` is constant. The simulation of the dataset can be done with
the function `sim_model`, which implements the simulation in two parts:

### Simulate the dataset

The data for our model can be simulated with the function `sim_model`.
The two main arguments of `sim_model` function, when working with linear Gaussian
models, are the `formula` and the sample size `n`.
In order to obtain a reproducible dataset, a seed must be defined
with the function `set.seed` or by using the argument `seed` in `sim_model`.

```{r}
data_model <- sim_model(formula = f, n = 1000, seed = 1, responses = 3)
```

The first 10 rows of the generated dataset looks as follows:
```{r, echo = FALSE, results = 'asis'}
knitr::kable(head(data_model, 10))
```

it contains an unique `id` for each individual, all the predictors included in the
`formula` (i.e. `x1` and `sex`), the parameters (`mean` and `sd`), and the simulated
`response` variable.

Some customization for the effects can be used, for instance the labels of the factor
can be included with the option `labels` in the function `fa`.
```{r}
library(ggplot2)
ggplot(data_model, aes(x, response)) +
  geom_point(aes(col = factor(response_label)), alpha = 1/5)

```

```{r}

Cor <- matrix(c(1, -0.9, 0, -0.9, 1, 0, 0, 0, 1), nrow = 3)
vars <- diag(c(1,1,3))
Sigma <- vars %*% Cor %*% vars

f <- list(
  # mean ~ mre(factor(id), sigma = get("Sigma")),
  mean ~ mfe(x, beta = c(-10, 0, 10)) + mre(factor(id), sigma = get("Sigma")),
  sd ~ I(0)
  )

data_model <- sim_model(formula = f, n = 1000, seed = 1, responses = 3)
knitr::kable(head(data_model, 10))
ggplot(data_model, aes(x, response)) +
  geom_point(aes(col = factor(response_label)), alpha = 1/5)

ggplot(data_model, aes(id, mre.factor.mean)) +
  geom_line(aes(col = factor(response_label)), alpha = 0.8) +
  facet_wrap(~ response_label, ncol = 1, scales = "free_y")

cor(matrix(data_model$mre.factor.mean, ncol = 3))

```

