---
title: "Data Simulation for Geostatistical Models"
author: "Erick A. Chacon-Montalvan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Simulation for Geostatistical Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
knitr::opts_chunk$set(comment = "#>")
options(width = 100)
```

In this vignette, we show how to simulate a geostatistical model.

#### Required packages

```{r}
library(datasim)
library(ggplot2)
```

## Geostatistical Model

First, we need to define a list of formulas specifying the type of effect that are
included in the linear predictor of each parameter. For example, this list can be
defined as follows.

```{r}
f <- list(
  mean ~ I(5) +
    gp(list(s1, s2), cor.model = "exp_cor", cor.params = list(phi = 0.02)),
  sd ~ I(1)
  )
```

In this formula, it can be seen that an intercept, a linear effect on `x1` and an
spatial Gaussian process in included on the `mean` parameter, while the
standard deviation `sd` is constant. The simulation of the dataset can be done with
the function `sim_model`.

### Simulate the dataset

The two main arguments of `sim_model` function, when working with linear Gaussian
models, are the `formula` and the sample size `n`.
In order to obtain a reproducible dataset, a seed must be defined
with the function `set.seed` or by using the argument `seed` in `sim_model`.

```{r}
data_model <- sim_model(formula = f, n = 500, seed = 1)
```

The first 10 rows of the generated dataset looks as follows:
```{r, echo = FALSE, results = 'asis'}
knitr::kable(head(data_model, 10))
```

it contains an unique `id` for each individual, the coordinates of the spatial effect (`s1` and `s2`), the realization of the Gaussian process `gp.list.mean`, the parameters (`mean` and `sd`) and the simulated `response` variable.

### Visualize

```{r}
ggplot(data_model, aes(s1, s2)) +
  geom_point(aes(col = gp.list.mean)) +
  scale_colour_distiller(palette = "RdYlBu")
```


### Simulate coordinates or predictors only

`sim_model` simulate the entire dataset. If only predictors want to be simulated to
have more control, the function `model_frame` can be used.  The two main arguments
of this function are the `formula` and the sample size `n`.

```{r}
data_frame <- model_frame(formula = f, n = 100, seed = 1)
```
The first 10 rows of the generated dataset looks as follows. As expected, only the
covariates are simulated.

```{r, echo = FALSE, results = 'asis'}
knitr::kable(head(data_frame, 10))
```

### Simulate the response variable only

If the data is already obtained, we can simulate the response variable using the
function `model_response`.

```{r}
data_frame <- model_response(data_frame, formula = f)
```

The first 10 rows of the generated dataset looks as follows. As expected, only the
covariates are simulated.

```{r, echo = FALSE, results = 'asis'}
knitr::kable(head(data_frame, 10))
```
